SDSS-IV MaNGA: The Radial Profile of Enhanced Star Formation in Close Galaxy Pairs

### 2020-09-22 maintex.tex
updated mass ratio and separation sections

\subsection{Dependency on Mass Ratio}
While the total stellar mass between the galaxies has no effect on the merger induced star formation, the mass ratio between the paired galaxies does. The mass ratio here is calculated as the difference between stellar masses contained within the 50\% half light radius,

\begin{equation}
\Delta {\rm log}(M) = {\rm log}(M_{target}) - {\rm log}(M_{comp}) 
\end{equation}

Where M$_{target}$ is the stellar mass of the MaNGA target galaxy and M$_{comp}$ is the stellar mass of the companion galaxy. Note that we do not use the absolute value of the difference like how the mass ratio is typically used. We do this so that we can independently study how the star formation large central galaxies behave in comparison to small satellite galaxies.

@@

In the Top row of Figure \ref{fig:ssfr_dmsep} we see that the sSFR enhancement is strongest in galaxies close to a 1:1 mass ratio, $\Delta$log(M) $=$ 0.0. These galaxies feature a central enhancement of $\sim$0.3 dex which is 0.1 dex higher than the average central enhancement of the whole sample. This enhancement falls with wider mass ratio; however, pairs with large differences in stellar mass still feature substantial levels of star formation enhancement, 0.15$-$0.20 dex at mass ratios of $|\Delta$log(M)$|$ $=$ 1.0.

We also use the mass ratio to explore how the merger induced tidal torques affect paired galaxies with large mass ratios, $|\Delta$log(M)$|$ > 0.5. We use the mass ratio to separate the massive central galaxy from satellite galaxies. A paired galaxy with a positive mass ratio is the more massive galaxy of the pair while a paired galaxy with negative mass ratio is the less massive galaxy. We see that the less massive galaxy of a pair show a higher level of sSFR enhancement, by $\sim$0.1 dex, with respect to the more massive galaxy. This shows us that satellite galaxies may be more susceptible to tidal interactions than the central galaxies in the pair system.

In previous studies it was noted that the enhancement to the SFR is strongest in galaxy pairs with small mass ratios \citet{Ellison:2008}. We see the same effect; however, we have also shown that galaxy pairs with large mass ratios still show a substantial level of sSFR enhancement. It has been a convention to split galaxy pairs in major mergers, $|$log(M)$|$ $\le$ 0.5, and minor mergers 0.5 $<$ $|$log(M)$|$ $\le$ 1.0. In this work, we can see that the mass ratio range used in galaxy pair studies could be widened to at least $|$log(M)$|$ $\le$ 1.5 so that larger pair samples can be constructed.

Further, within these galaxy pairs with wide mass ratios we see that the less massive galaxy in the pair shows higher levels of sSFR enhancement. This implies that satellite galaxies are more affected by interaction induced star formation than the central galaxy.

@@

We also look at the level of sSFR enhancement as a function of the projected separation between the two galaxies in the bottom row of Figure \ref{fig:ssfr_dmsep}. The $\Delta$log(sSFR) profiles show a clear gradient with the projected separation where close galaxy pairs show higher levels of sSFR enhancement compared to pairs with wide separations. The profiles in the galaxy pairs in the separation range, $r_p$ $=$ 0.0$-$12.5, are generally $\sim$0.1 dex above the median and the the profiles of the galaxy pairs in the separation range, $r_p$ $=$ 37.5$-$50.0, are generally $\sim$0.1 dex below the median.

@@

The sSFR enhancement increasing with closer projected separations is to be expected and has been shown in previous studies \citep{Li:2008, Ellison:2008, Scudder:2012, Patton:2013}. The passage of the first pericenter is predicted to trigger a burst of star formation which falls as the two galaxies separate \citep{Scudder:2012}. The sSFR enhancement in our sample persists out to 50 kpc, \citet{Patton:2013} showed that the burst of star formation can persist out to 150 kpc.

For galaxies with separations of $r_p$ $<$ 5 kpc and $r_p$ $=$ 30$-$35 kpc we find a $\Delta$log(sSFR) of zero. For the $r_p$ $<$ 5 kpc galaxies, about half of the pairs show signs of tidal features while the others do not. The pairs without the tidal features may be approaching their first pericenter, meaning that the merger-induced gas-inflows have not been triggered yet. The dips in $\Delta$log(sSFR) may also just be due to statistical error from the relatively limited sample size.

Between the three parameters, stellar mass, mass ratio, and projected separation, the projected separation has the strongest effect on the $\Delta$log(sSFR) enhancement, followed by the mass ratio. From this we see that the paired galaxies with the strongest enhancements to the central sSFR are paired galaxies with 1:1 mass ratios and  which are within 10 kpc.

Between the splitting the $\Delta$log(sSFR) profile by stellar mass, mass ratio, and projected separation, we see that the general shape of the profile within the inner 1.0 R/\reff\ is unchanged. The profile is peaked in the center of the paired galaxies and linearly (in log space) to zero around 1.0 R/\reff. Different projected separations and mass ratios may elevate or suppress the level of the enhancement in the profile's center; however, the shape of the profile is largely unchanged.

@@

We primarily focused on how the $\Delta$log(sSFR) in paired galaxies behaves as a function of stellar mass, but we also explored how the $\Delta$log(sSFR) behaves as a function of the mass ratio between the paired galaxies and the projected separation between the paired galaxies. While we see hints that the $\Delta$log(sSFR) is affected by the mass ratio and projected separation, these results are less definitive than the $\Delta$log(sSFR) profile's lack of dependence on stellar mass.

>>>>

\subsection{Dependency on Mass Ratio}
While the total stellar mass of the galaxies has no effect on the merger induced star formation, the mass ratio between the paired galaxies does. The mass ratio here is calculated as the difference between stellar masses contained within the 50\% half light radius,

\begin{equation}
\Delta {\rm log}(M) = {\rm log}(M_{\rm target}) - {\rm log}(M_{\rm comp}) 
\end{equation}

Where $M_{\rm target}$ is the stellar mass of the MaNGA target galaxy and $M_{\rm comp}$ is the stellar mass of the companion galaxy.

@@

In the Top row of Figure \ref{fig:ssfr_dmsep} we see that the sSFR enhancement is strongest in galaxies close to a 1:1 mass ratio, $\Delta$log($M$) $=$ 0.0. These galaxies feature a central enhancement of $\sim$0.3 dex which is 0.1 dex higher than the average central enhancement of the whole sample. This enhancement falls with wider mass ratio; however, pairs with large differences in stellar mass still feature substantial levels of star formation enhancement, 0.15$-$0.20 dex at mass ratios of $|\Delta$log($M$)$|$ $=$ 1.0.

We also use the mass ratio to explore how the merger induced tidal torques affect paired galaxies with large mass ratios, $|\Delta$log($M$)$|$ > 0.5. We use the mass ratio to separate the massive central galaxy from satellite galaxies. A paired galaxy with a positive mass ratio is the more massive galaxy of the pair while a paired galaxy with negative mass ratio is the less massive galaxy. We see that the less massive galaxy of a pair show a higher level of sSFR enhancement, by $\sim$0.1 dex, with respect to the more massive galaxy. This shows us that satellite galaxies may be more susceptible to tidal interactions than the central galaxies in the pair system.

In previous studies it was noted that the enhancement to the SFR is strongest in galaxy pairs with small mass ratios \citet{Ellison:2008}. We see the same effect; however, we have also found that galaxy pairs with large mass ratios still show a substantial level of sSFR enhancement. It has been a convention to split galaxy pairs in major mergers, $|$log($M$)$|$ $\le$ 0.5, and minor mergers 0.5 $<$ $|$log($M$)$|$ $\le$ 1.0. In this work, we can see that the mass ratio range used in galaxy pair studies could be widened to at least $|$log($M$)$|$ $\le$ 1.5 so that larger pair samples can be constructed.

Further, within these galaxy pairs with wide mass ratios we see that the less massive galaxy in the pair ($\Delta$log($M$) $<$ 0) shows higher levels of sSFR enhancement. This implies that satellite galaxies are more affected by interaction induced star formation than the central galaxy.

@@

We also look at the level of sSFR enhancement as a function of the projected separation between the two galaxies in the bottom row of Figure \ref{fig:ssfr_dmsep}. The $\Delta$log(sSFR) profiles show a clear gradient with the projected separation where close galaxy pairs show higher levels of sSFR enhancement compared to pairs with wide separations. The profiles in the galaxy pairs in the separation range, $r_{\rm p}$ $=$ 0.0$-$12.5, are generally $\sim$0.1 dex above the median and the profiles of the galaxy pairs in the separation range, $r_{\rm p}$ $=$ 37.5$-$50.0, are generally $\sim$0.1 dex below the median.

@@

The sSFR enhancement increasing with closer projected separations is to be expected and has been shown in previous studies \citep{Li:2008, Ellison:2008, Scudder:2012, Patton:2013}. The passage of the first pericenter is predicted to trigger a burst of star formation which falls as the two galaxies separate \citep{Scudder:2012}. The sSFR enhancement in our sample persists out to 50 kpc. \citet{Patton:2013} showed that the burst of star formation can persist out to 150 kpc.

For galaxies with separations of $r_{\rm p}$ $<$ 5 kpc and $r_{\rm p}$ $=$ 30$-$35 kpc we find a $\Delta$log(sSFR) of zero. For the $r_{\rm p}$ $<$ 5 kpc galaxies, about half of the pairs show signs of tidal features while the others do not. The pairs without the tidal features may be approaching their first pericenter, meaning that the merger-induced gas-inflows have not been triggered yet. The dips in $\Delta$log(sSFR) may also just be due to statistical error from the relatively limited sample size.

@@

Among the three parameters, stellar mass, mass ratio, and projected separation, the projected separation has the strongest effect on the $\Delta$log(sSFR) enhancement, followed by the mass ratio. From this we see that the paired galaxies with the strongest enhancements to the central sSFR are paired galaxies with 1:1 mass ratios and  which are within 10 kpc.

Between the splitting the $\Delta$log(sSFR) profile by stellar mass, mass ratio, and projected separation, we see that the general shape of the profile within the inner 1.0 $R$/\reff\ is unchanged. The profile is peaked in the center of the paired galaxies and linearly (in log space) to zero around 1.0 $R$/\reff. Different projected separations and mass ratios may elevate or suppress the level of the enhancement in the profile's center; however, the shape of the profile is largely unchanged.
